Trail Saving on Backtrack

### 2020-02-28 paper.tex
plot

\begin{document}
\maketitle
\begin{abstract}
    A CDCL \sat solver can backtrack a large distance when it learns a
    new clause, e.g, when the new learnt clause is a unit clause the
    solver has to backtrack to level zero. When the length of the
    backtrack is large, the solver can end up reproducing many of the
    same decisions and propagations when it redescends the search
    tree. Motivated by this potential redundancy, partial backtracking
    (chronological backtracking) has been proposed. This technique
    allows the solver to reduce the length of its backtrack, e.g.,
    allowing the solver to backtrack to the previous level after a
    conflict is found. However, this technique has its shortcomings
    and only tends to work when limits are placed on its
    application. In this paper we present a new trail saving technique
    that makes a copy of the part of the trail that was backtracked
    over. This saved copy can then be used to improve the efficiency
    of the solver's subsequent redescent. Although our technique does
    not save as much work as chronological backtracking, it also does
    not have chronological backtracking's shortcomings. Furthermore,
    the saved trail also provides the solver with the ability to look
    ahead along the previous trail. This can be exploited to improve
    its efficiency. Our empirical results show that our technique
    often yields superior performance over chronological backtracking,
    and that is is able to improve the performance of state of the art
    solvers.

@@

\begin{figure} \centering \includegraphics[scale=0.7]{figures/scatter.pdf}\caption{Comparing propagations per second with and without trail saving in Cadical.}\label{fig:props}\end{figure}

\section{Conclusion}
We have shown that our method trail saving can speed up two
state-of-the-art \sat solvers, cadical and MapleSAT, about as well as
chronological backtracking can without breaking the conventional invariants
of CDCL \sat solvers. We also introduced three enhancements one can
implement if they are using a saved trail and demonstrated experimentally
that these enhancements can sometimes improve the performance by a significant
amount. Future work could include tuning the parameters of the enhancements 
to work optimally or introducing further enhancements, such as using the saved
trail to help make inprocessing techniques faster or
using the saved trail to learn a second learnt clause from a single conflict.
Lastly, it might be possible to combine trail saving with chronological
backtracking and figure out when each is most beneficial.

>>>>

\begin{document}
\maketitle
\begin{abstract}
    A CDCL \sat solver can backtrack a large distance when it learns a
    new clause, e.g, when the new learnt clause is a unit clause the
    solver has to backtrack to level zero. When the length of the
    backtrack is large, the solver can end up reproducing many of the
    same decisions and propagations when it redescends the search
    tree. Motivated by this potential redundancy, partial backtracking
    (chronological backtracking) has been proposed. This technique
    allows the solver to reduce the length of its backtrack, e.g.,
    allowing the solver to backtrack to the previous level after a
    conflict is found. However, this technique has its shortcomings
    and only tends to work when limits are placed on its
    application. In this paper we present a new trail saving technique
    that makes a copy of the part of the trail that was backtracked
    over. This saved copy can then be used to improve the efficiency
    of the solver's subsequent redescent. Although our technique does
    not save as much work as chronological backtracking, it also does
    not have chronological backtracking's shortcomings. Furthermore,
    the saved trail also provides the solver with the ability to look
    ahead along the previous trail. This can be exploited to improve
    its efficiency. Our empirical results show that our technique
    often yields superior performance over chronological backtracking,
    and that is is able to improve the performance of state-of-the-art
    solvers.

@@

\begin{figure}\centering\includegraphics[scale=0.65]{figures/cactus_cadical_new.pdf}\caption{\small{Cactus plot for the newest version of cadical comparing standard non-chronological backtracking to \cbt and various configurations of trail saving. The first 400 problems were solved in less than 1200 seconds, so that part of the plot is truncated.}}\label{fig:cactus_cadical}\end{figure}

\begin{figure}\centering\includegraphics[scale=0.7]{figures/scatter.pdf}\caption{Comparing propagations per second with and without trail saving in Cadical.}\label{fig:props}\end{figure}

\section{Conclusion}
We have shown that our method trail saving can speed up two
state-of-the-art \sat solvers, cadical and MapleSAT, about as well as
chronological backtracking can without breaking the conventional invariants
of CDCL. We also introduced three enhancements one can
implement when using a saved trail and demonstrated experimentally
that these enhancements can sometimes improve a solver's performance by a significant
amount. We have shown that trail saving and all enhancements we proposed are sound.
A line of future work might include introducing further enhancements,
such as using the saved trail to help make inprocessing techniques faster or
using the saved trail to learn a second learnt clause from a single conflict.
It is also possible to combine trail saving with chronological
backtracking if one is willing to break the traditional invariants of CDCL,
but we leave it to future work to determine whether or not this would be
useful and how to best approach it.
